---
title: "PoSI Case Study: Telomere Length Analysis"
author: ''
output:
  html_document:
    code_folding: show
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: true
  pdf_document:
    toc: yes
    toc_depth: '2'
bibliography: bibliography.bib
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "fig/")
if (!require("pacman")) install.packages(pacman)
pacman::p_load(data.table, sandwich, jtools, xtable, scales, dplyr, ggplot2, devtools, matrixStats, knitr, kableExtra)

if (!require("tmax")) devtools::install_github(repo="post-selection-inference/R", subdir = "tmax")
library(tmax)
```


This file demonstrates PoSI on the Telomere Length (TL) analysis by @nersisyan2019wgs. The goal of the analysis is to study the inheritance patterns and associated genetic factors. The data and scripts of their analyses are available on [Github](https://github.com/lilit-nersisyan/gonl_tqtl). Here we focus on the first part of the multiple linear regression (MLR) analysis to understand the TL inhereitance patterns.

# MTL data

There are in total 250 family trios in the original data. According to the article, two of the families with missing data for the mother were removed, and two families with discordant age differences at the time of data collection and at conception were also discarded. Hence there are 246 samples left in the `child.df` data. The description of the data is as follows.

Variable       |  Description  
-------------  |--------------------------------------------------------------------
GoNL_Sample_ID | id of Genome of the Netherlands (GoNL) project 
family         | family ID
member         | family member (here are all children, `c`)
Sex            | sex
Age            | age
MTL            | mean telomere length
ageC           | NA
MAC            | mother's age at conception
PAC            | father's age at conception
mMTL           | mother's mean telomere length
fMTL           | father's mean telomere length
mAge           | mother's age
fAge           | father's age

```{r read data}
# phenotype data for children and parents 246 families
load("gonl_tqtl-master/child.df.RData") 

# final model
fit.final <- lm(MTL ~ Age + mMTL + fMTL + MAC + PAC, child.df)

# summary table of the final model
summary(fit.final)
# summary table of the final model with sandwich
summ(fit.final, robust = "HC0")
```


```{r prep data}
# fit.0 <- lm(MTL ~ (Sex + Age + mMTL + fMTL + MAC + PAC)^2, child.df)
x <- model.matrix(MTL ~ (Sex + Age + mMTL + fMTL + MAC + PAC)^2, child.df)
y <- child.df$MTL

# final model
M.final <- 3:7
colnames(x)[M.final]

p <- ncol(x) - 1
nboot <- 1000
```

# PoSI

We provide PoSI adjustment considering selection mentioned in @nersisyan2019wgs.

* Variables: `Sex`, `Age`, `mMTL`, `fMTL`, `MAC`, and `PAC`;
* Interactions: All pairwise interactions.

```{r posi, eval=F}
Hm <- matrix(NA, nrow = p, ncol = nboot)
Hs <- matrix(NA, nrow = p, ncol = nboot)

Hm1 <- matrix(NA, nrow = p, ncol = nboot)
Hs1 <- matrix(NA, nrow = p, ncol = nboot)

Hmp <- array(NA, c(p, nboot, p))
Hsp <- array(NA, c(p, nboot, p))


ks <- 1:p
for(i in seq_along(ks)) {
  k <- ks[i]
  print(k)
  tmp <- max_t_mul_boot_k(x[,-1], y, k,
                          sandwich = T, return_sample = T, 
                          Nboot = nboot, intercept = T,
                          individual = 1:p)
  Hs[i, ] <- colMax(tmp$BootSample)
  Hm[i, ] <- colMax(tmp$BootRank)
  
  # Hs1[i, ] <- colMax(tmp$BootSample1)
  # Hm1[i, ] <- colMax(tmp$BootRank1)
  Hsp[i,,] <- t(tmp$BootSample1)
  Hmp[i,,] <- t(tmp$BootRank1)
  
}

# PoSI constant with prepivot
K_norm <- get_T(x[,-1], y, M.final, NULL, Hm, maxk = p, Nboot = nboot, intercept = T, adjust = T)
# PoSI1 constant with prepivot
K_norm1 <- sapply(M.final, function(i) get_T(x[,-1], y, M.final, NULL, Hmp[,,i], maxk = p, Nboot = nboot, intercept = T, adjust = T))

# adjusted p
se <- diag(sqrt(vcovHC(fit.final, "HC0")))[-1]
t.s <- coef(fit.final)[-1]/se
pval_norm <- get_p(t.s, x[,-1], y, M.final, NULL, Hm, maxk = p, Nboot = nboot, intercept = T, adjust = T)
pval_norm1 <- sapply(1:length(M.final), function(i) get_p(t.s[i], x[,-1], y, M.final, NULL, Hmp[,,M.final[i]], maxk = p, Nboot = nboot, intercept = T, adjust = T))

saveRDS(list(K_norm = K_norm, K_norm1 = K_norm1,
             pval_norm = pval_norm, pval_norm1 = pval_norm1),
        "MTL_PoSI.RDS")
```

# Report

After PoSI or PoSI1 adjustment, `AGE` and `MAC` change from siginificant at 0.05 level to insignificant.


Variable       |  Unadjusted | Adjusted  
-------------  | ------- | -------
Age            |  &#10003; | &#10007;
mMTL           | &#10003; | &#10003;
fMTL           | &#10003; | &#10003;
MAC            | &#10003; | &#10007;
PAC            | &#10007; | &#10007;


```{r posi table}
ret <- readRDS("MTL_PoSI.RDS")
K_norm <- ret$K_norm; K_norm1 <- ret$K_norm1
pval_norm <- ret$pval_norm; pval_norm1 <- ret$pval_norm1

fit.final <- lm(MTL ~ Age + mMTL + fMTL + MAC + PAC, child.df)

se <- diag(sqrt(vcovHC(fit.final, "HC0")))[-1]
fit.final.ci <- summ(fit.final, confint = T)$coeftable[-1,]

ret.p <- 
  data.table(var = factor(colnames(x)[M.final], 
                          levels = colnames(x)[M.final]),
             pval_t = summ(fit.final)$coeftable[-1, 4],
             pval_adjust = pval_norm,
             pval_adjust1 = pval_norm1)

ret.ci <- 
  data.table(var = factor(rep(colnames(x)[M.final], 3),
                          levels = colnames(x)[M.final]),
             method = factor(rep(c("t", "adjust", "adjust1"), 
                                 each = length(M.final)),
                             levels = c("t", "adjust", "adjust1")),
             lower = c(fit.final.ci[,2], 
                       fit.final$coefficients[-1] - K_norm * se,
                       fit.final$coefficients[-1] - K_norm1 * se
             ),
             upper = c(fit.final.ci[,3], 
                       fit.final$coefficients[-1] + K_norm * se,
                       fit.final$coefficients[-1] + K_norm1 * se
             )
  )

# xtable(ret.ci)
ret.ci.wide <- dcast(ret.ci, var ~ method, value.var = c("lower", "upper"))
name.order <- 1
for(method in c("t", "adjust", "adjust1")) {
  col.idx <- grep(paste0("^", method, "$"), 
                  sapply(strsplit(names(ret.ci.wide), "_"), tail, 1))
  ret.ci.wide[, eval(method)] <- 
    Reduce("*", 
           ret.ci.wide[, col.idx, 
                       with=F]) > 0
  name.order <- c(name.order, col.idx, ncol(ret.ci.wide))
}
ret.ci.wide <- ret.ci.wide[,name.order,with=F]

ret.ci.wide <- dcast(ret.ci, var ~ method, value.var = c("lower", "upper"))
name.order <- 1
for(method in c("t","adjust", "adjust1")) {
  col.idx <- grep(paste0("^", method, "$"), 
                  sapply(strsplit(names(ret.ci.wide), "_"), tail, 1))
  ret.ci.wide[, eval(method)] <- 
    Reduce("*", 
           ret.ci.wide[, col.idx, 
                       with=F]) > 0
  # name.order <- c(name.order, col.idx, ncol(ret.ci.wide))
}
ret.ci.wide$K_t <- 1.96
ret.ci.wide$K_adjust <- K_norm
ret.ci.wide$K_adjust1 <- K_norm1
ret.ci.wide <- merge(ret.ci.wide, ret.p, by = "var")

for(method in c("t","adjust", "adjust1")) {
  col.idx <- grep(paste0("^", method, "$"), 
                  sapply(strsplit(names(ret.ci.wide), "_"), tail, 1))
  name.order <- c(name.order, col.idx)
}
ret.ci.wide <- ret.ci.wide[,name.order,with=F]

changed.idx <- as.matrix(ret.ci.wide[, .SD, .SDcols = c("t", "adjust1")])
changed.idx[changed.idx] <- 1; changed.idx[!changed.idx] <- -1
changed.idx <- which(rowProds(changed.idx) != 1)

# the unadjusted with sandwich
kable(ret.ci.wide[, 1:6]) %>% 
  kable_styling("striped", full_width = F) %>%
  row_spec(changed.idx, bold = T, color = "white", background = "#D7261E")
# PoSI adjusted 
kable(ret.ci.wide[, c(1, 7:11)]) %>% 
  kable_styling("striped", full_width = F) %>%
  row_spec(changed.idx, bold = T, color = "white", background = "#D7261E")
# PoSI1 adjusted
kable(ret.ci.wide[, c(1, 12:16)]) %>% 
  kable_styling("striped", full_width = F) %>%
  row_spec(changed.idx, bold = T, color = "white", background = "#D7261E")
```


```{r plot}
ret.ci %>%
  filter(method %in% c("t", "berk", "adjust1")) %>%
  ggplot(aes(x = var, col = method)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), position=position_dodge(0.6)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_manual(values = hue_pal()(4),
                     labels = c("Naive", "PoSI", "Adjusted1")) +
  xlab("") +
  ylab("") + 
  theme(legend.key=element_blank(),
        legend.key.width = unit(1,"cm"),
        legend.background=element_blank(),
        legend.position = c(1, 0.85),
        legend.direction = "vertical",
        legend.justification="right",
        legend.spacing.x = unit(0.05, 'in'),
        legend.title = element_blank(), 
        legend.text = element_text(size = 14),
        axis.text.x = element_text(size = 14))
```