---
title: "PoSI Case Study: Boston Housing"
author: ''
output:
  html_document:
    code_folding: show
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: true
  pdf_document:
    toc: yes
    toc_depth: '2'
bibliography: bibliography.bib
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "fig/")
if (!require("pacman")) install.packages(pacman)
pacman::p_load(data.table, sandwich, jtools, xtable, scales, dplyr, ggplot2, devtools, matrixStats)

if (!require("tmax")) devtools::install_github(repo="post-selection-inference/R", subdir = "tmax")
library(tmax)
```


This page demonstrate the PoSI on the Boston Housing data from @harrison1978hedonic with amendments by @gilley1996harrison.

# The Boston housing data
The data description is as follows.


Variable     |  Description  
-------------|--------------------------------------------------------------------
 CRIM        |  crime rate  
 ZN          |  proportion of residential land zoned for lots over 25,000 sq. ft 
 INDUS       |  proportion of non-retail business acres 
 CHAS        |  Charles River dummy variable (=1 if tract bounds river; 0 otherwise) 
 NOX         |  nitric oxides concentration, pphm 
 RM          |  average number of rooms per dwelling 
 AGE         |  proportion of owner-occupied units built prior to 1940 
 DIS         |  weighted distances to five Boston employment centers 
 RAD         |  index of accessibility to radial highways 
 TAX         |  full-value property tax rate per $10,000 
 PTRATIO     |  pupil teacher ratio 
 B           |  1000*(Bk-0.63)^2 where Bk is the proportion of blacks
 LSTAT       |  percent lower status population 
 MEDV        |  median value of owner occupied homes in $1000's 

```{r aux}
box.cox <- function(y, lambda=0) {
  if(lambda==0) log(y) else (y^lambda-1)/(lambda-1) }

power.trans <- function(y, lambda = 0) {
  if(lambda == 0) log(y) else y^lambda
}
```

```{r read data}
data <- read.table("boston.dat")

# https://spatial-statistics.com/pace_manuscripts/jeem_ms_dir/pdf/fin_jeem.pdf
data$MEDV[c(8,39,191,241,438,443,455,506)] <-
  c(22.1, 24.2, 33, 27, 8.2, 14.8, 14.4, 19)
data$LSTAT <- data$LSTAT/100

# the final model
fit <- lm(log(MEDV) ~ 
            I(RM^2) + 
            AGE +
            I(log(DIS)) + I(log(RAD)) +
            TAX +
            PTRATIO +
            B +
            I(log(LSTAT)) +
            CRIM + ZN + INDUS + CHAS +
            I(box.cox(NOX, 2)),
          data)
summ(fit)
summary(fit)
```

# PoSI

We provide PoSI adjustment considering selection of

* Log transformation of the response `MEDV`
* Grid search of Box-Cox transformation of the variable of interest `NOX`
* Log/linear/quadratic transformation of `RM`

```{r prep data}
x <- data[, 1:(ncol(data)-1)]
y <- data[, ncol(data)]

log_y <- log(y)
y_ <- cbind(y, log_y)

x$DIS <- log(x$DIS)
x$RAD <- log(x$RAD)
x$LSTAT <- log(x$LSTAT)

p <- ncol(x)
nboot <- 1000
```


```{r posi}
# grid search from 0 to 10 with 1000 grid points
nox_trans <- seq(0,10,length.out = 1000)
# log/linear/quadratic transformation for RM
rm_trans <- 0:2
combs <- t(expand.grid(nox_trans, rm_trans))

Hm <- matrix(0, nrow = ncol(combs), ncol = nboot)
Hs <- matrix(0, nrow = ncol(combs), ncol = nboot)

# for each variable, (maxk, nboot)
Hmp <- array(0, c(ncol(combs), nboot, p))
Hsp <- array(0, c(ncol(combs), nboot, p))

for(j in ncol(y_)) {
  for(i in 1:ncol(combs)) {
    if((i %% 500)==0) print(c(j,i))
    q <- combs[,i]
    x_ <- x
    x_$NOX <- box.cox(x$NOX, q[1])
    x_$RM <- power.trans(x$RM, q[2])
    
    tmp <- max_t_mul_boot_k(x_, y_[,j], p,
                            sandwich = T,
                            intercept = T, 
                            Nboot = nboot,
                            force_in = 1:p,
                            individual = 1:p)
    
    
    
    Hs[i, ] <- pmax(Hs[i, ], colMax(tmp$BootSample))
    Hm[i, ] <- pmax(Hm[i, ], colMax(tmp$BootRank))
    
    Hsp[i,,] <- pmax(Hsp[i,,], t(tmp$BootSample1))
    Hmp[i,,] <- pmax(Hmp[i,,], t(tmp$BootRank1))
  }  
}

# saveRDS(list(Hm = Hm, Hs = Hs, Hmp = Hmp, Hsp = Hsp), "boston.rds")
```

```{r posi table}
ret <- readRDS("boston.rds")
Hm <- ret$Hm; Hs <- ret$Hs; Hmp <- ret$Hmp; Hsp <- ret$Hsp

Hm_ <- colMax(Hm)

x_ <- x
x_$NOX <- box.cox(x$NOX, 2)
# x_$NOX <- power.trans(x$NOX, 2)
x_$RM <- power.trans(x$RM, 2)

# PoSI constant with prepivot
K <- get_T(Hm = Hs)
# PoSI constant without prepivot
K_norm <- get_T(x_, y_[,2], 1:p, 1, matrix(Hm_, nrow = 1), maxk = 1, Nboot = nboot, intercept = T, adjust = T)
# PoSI1 constant with prepivot
K_norm1 <- sapply(1:p, function(i) get_T(x_, y_[,2], 1:p, 1, matrix(colMax(Hmp[,,i]),nrow=1), maxk = 1, Nboot = nboot, intercept = T, adjust = T))


fit <- lm(log_y~as.matrix(x_))
se <- diag(sqrt(vcovHC(fit, "HC0")))[-1]
fit.ci <- summ(fit, confint = T)$coeftable[-1,]

# adjusted p
t.s <- coef(fit)[-1]/se
pval_norm <- get_p(t.s, x_, y_[,2], 1:p, 1, matrix(Hm_, nrow = 1), maxk = 1, Nboot = nboot, intercept = T, adjust = T)
pval_norm1 <- sapply(1:p, function(i) get_p(t.s[i], x_, y_[,2], 1:p, 1, matrix(colMax(Hmp[,,i]),nrow=1), maxk = 1, Nboot = nboot, intercept = T, adjust = T))

colnames(x) <- c("CRIM", "RES", "INDUS", "RIVER", "NOX", 
                 "RM", "AGE", "DEMPC", "DRADH", "TAX", 
                 "PTR", "BLK", "LSTAT")
colname_levels <- c("NOX", "RM", "AGE", "CRIM", "RES", "INDUS",
                    "RIVER", "TAX", "PTR", "BLK", "LSTAT", "DEMPC", "DRADH")
colname_labels <- c("NOX^2", "RM^2", "AGE", "CRIM", "RES", "INDUS",
                    "RIVER", "TAX", "PTR", "BLK", "logLSTAT", "logDEMPC", "logDRADH")

ret.p <- 
  data.table(var = factor(colnames(x),
                          levels = colname_levels),
             pval_t = summ(fit)$coeftable[-1, 4],
             pval_adjust = pval_norm,
             pval_adjust1 = pval_norm1)
ret.ci <- 
  data.table(var = factor(rep(colnames(x), 3),
                          levels = colname_levels),
             method = factor(rep(c("t", "adjust", "adjust1"), 
                                 each = p),
                             levels = c("t", "adjust", "adjust1")),
             lower = c(fit.ci[,2], 
                       # fit$coefficients[-1] - posi.k1 * se,
                       fit$coefficients[-1] - K_norm * se,
                       fit$coefficients[-1] - K_norm1 * se
             ),
             upper = c(fit.ci[,3], 
                       # fit$coefficients[-1] + posi.k1 * se,
                       fit$coefficients[-1] + K_norm * se,
                       fit$coefficients[-1] + K_norm1 * se
             )
  )

xtable(ret.ci)

ret.ci.wide <- dcast(ret.ci, var ~ method, value.var = c("lower", "upper"))
name.order <- 1
for(method in c("t","adjust", "adjust1")) {
  col.idx <- grep(paste0("^", method, "$"), 
                  sapply(strsplit(names(ret.ci.wide), "_"), tail, 1))
  ret.ci.wide[, eval(method)] <- 
    Reduce("*", 
           ret.ci.wide[, col.idx, 
                       with=F]) > 0
  # name.order <- c(name.order, col.idx, ncol(ret.ci.wide))
}
ret.ci.wide$K_t <- 1.96
ret.ci.wide$K_adjust <- K_norm
ret.ci.wide$K_adjust1 <- K_norm1
ret.ci.wide <- merge(ret.ci.wide, ret.p, by = "var")

for(method in c("t","adjust", "adjust1")) {
  col.idx <- grep(paste0("^", method, "$"), 
                  sapply(strsplit(names(ret.ci.wide), "_"), tail, 1))
  name.order <- c(name.order, col.idx)
}
ret.ci.wide <- ret.ci.wide[,name.order,with=F]
```


```{r plot}
ret.ci %>%
  filter(method %in% c("t", "adjust1")) %>%
  ggplot(aes(x = var, col = method)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), position=position_dodge(0.6)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_manual(values = hue_pal()(3),
                     labels = c("Unadjusted", "Adjusted")) +
  scale_x_discrete(labels = colname_labels) +
  xlab("") +
  ylab("") + 
  theme(legend.key=element_blank(),
        legend.key.width = unit(1,"cm"),
        legend.background=element_blank(),
        # legend.position = c(0.02, 0.10),
        legend.position = c(0.75, 0.18),
        legend.direction = "vertical",
        legend.justification="left",
        legend.spacing.x = unit(0.05, 'in'),
        legend.title = element_blank(), 
        legend.text = element_text(size = 14),
        axis.text.x = element_text(size = 14, angle = -30, hjust = -.01),
        plot.margin = margin(t = 5, b = -10, r = 20)) +
  coord_cartesian(xlim = c(.5,13.5))


ret.ci %>%
  filter(method %in% c("t", "adjust1")) %>%
  ggplot(aes(x = var, col = method)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), position=position_dodge(0.6)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_manual(values = hue_pal()(3),
                     labels = c("Unadjusted", "Adjusted")) +
  scale_x_discrete(labels = colname_labels) +
  xlab("") +
  ylab("") + 
  theme(legend.key=element_blank(),
        legend.key.width = unit(1,"cm"),
        legend.background=element_blank(),
        legend.position = c(0.75, 0.18),
        legend.direction = "vertical",
        legend.justification="left",
        legend.spacing.x = unit(0.05, 'in'),
        legend.title = element_blank(), 
        legend.text = element_text(size = 14),
        axis.text.x = element_text(size = 14, 
                                   angle = -30,
                                   hjust = -.01),
        plot.margin = margin(t = 5, b = -10, r = 20)) +
  coord_cartesian(xlim = c(.5,13.5), ylim=c(-.01, .01))

```
